Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2021-06-07 doc/manuscript/ESSDmanuscript.Rmd
rough start at intro, fill in author affiliations

affiliation:
  - code: 1
    address: Center for Conservatiton Ecology, Smithsonian Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: 
  - code: 3
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 4
    address: 
  - code: 5
    address:  IPCC Task Force on National Greenhouse Gas Inventories Technical Support Unit, Institute for Global Environmental Strategies, Hayama, Japan
  - code: 6
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 7
    address: CIRAD, Montpellier, France

@@

**(Forest are critical for climate change mitigation)**

**(Need for good data in international carbon accounting)**

**(Introduce EFDB & ForC)**

Here, we:
(1) clarify definitions of relevant carbon stocks and increments
(2) describe mapping of ForC to IPCC's EFDB,
(3) describe updates to ForC  (ForC v4.0),
(4) summarize the data in ForC that's relevant to EFDB, identifying gaps, and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

>>>>

affiliation:
  - code: 1
    address: Center for Conservatiton Ecology, Smithsonian Conservation Biology Institute, Front Royal, VA, United States
  - code: 2
    address: School of Geography, University of Leeds, Leeds, UK
  - code: 3
    address: Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States
  - code: 4
    address: The Nature Conservancy; Arlington VA 22203, USA
  - code: 5
    address:  IPCC Task Force on National Greenhouse Gas Inventories Technical Support Unit, Institute for Global Environmental Strategies, Hayama, Japan
  - code: 6
    address: Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama
  - code: 7
    address: CIRAD, Montpellier, France

@@

**As we seek to mitigate climate change [@unfccc_adoption_2015], forests are critical.**
*(See intros to other papers for relevant refs)*
Under Paris Agreement, **##**% of Nationally Determined Contributions relate to forests [@grassi_key_2017].

**The International Panel on Climate Change (IPCC) provides guidance for national greenhouse gas inventories [(REFS for older guidelines), @ipcc_2019_2019].**
These include guidance for accounting for forest land [@ipcc_agriculture_2006; @ipcc_2019_2019a].
This guidance has improved as more of the relevant underlying data has become available [@requenasuarez_estimating_2019], but there remains room for continuous improvement as the science advances.
For example, @cook-patton_mapping_2020 *(summarize IPCC comparison results)*. 
*Moreover, it is useful for those compiling national greenhouse gas inventories to have access to locally-specific information, when available.*
To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.

**The Global Forest Carbon Database, ForC, is the largest collection of published estimates of forest carbon stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021].**
*(add stats/ details)*
As such, ForC is positioned to improve forest C accounting through the transfer of data to EFDB.
The purpose of this publication is to document that process and provide recommendations for future improvements.

**Here, we**
(1) clarify definitions of relevant carbon stocks and increments
(2) describe mapping of ForC to IPCC's EFDB,
(3) describe updates to ForC  (ForC v4.0),
(4) summarize the data in ForC that's relevant to EFDB, identifying gaps, and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

### 2021-06-07 doc/manuscript/ESSDmanuscript.tex
rough start at intro, fill in author affiliations

\Author[1]{Madison}{Williams}
\Author[1]{Valentine}{Herrmann}
\Author[1,2]{Rebecca}{Banbury Morgan}
\Author[3]{Ben}{Bond-Lamberty (confirm)}
\Author[4]{Susan}{Cook-Patton}
\Author[5]{Sandro}{Federici}
\Author[6]{Helene}{Muller-Landau}
\Author[7]{Camille}{Piponiot}
\Author[1]{Teagan}{Rogers}
\Author[5]{Valentyna}{Slivinska}
\Author[1,6 *]{Kristina J.}{Anderson-Teixeira}

\affil[1]{Center for Conservatiton Ecology, Smithsonian Conservation
Biology Institute, Front Royal, VA, United States}
\affil[2]{}
\affil[3]{}
\affil[4]{}
\affil[5]{IPCC Task Force on National Greenhouse Gas Inventories
Technical Support Unit, Institute for Global Environmental Strategies,
Hayama, Japan}
\affil[6]{Forest Global Earth Observatory, Smithsonian Tropical Research
Institute, Panama, Republic of Panama}
\affil[7]{CIRAD, Montpellier, France}

@@

\textbf{(Forest are critical for climate change mitigation)}

\textbf{(Need for good data in international carbon accounting)}

\textbf{(Introduce EFDB \& ForC)}

Here, we: (1) clarify definitions of relevant carbon stocks and
increments (2) describe mapping of ForC to IPCC's EFDB, (3) describe
updates to ForC (ForC v4.0), (4) summarize the data in ForC that's
relevant to EFDB, identifying gaps, and (5) provide recommendations for
improving data collection, analysis, database, and accounting.

>>>>

\Author[1]{Madison}{Williams}
\Author[1]{Valentine}{Herrmann}
\Author[1,2]{Rebecca}{Banbury Morgan}
\Author[3]{Ben}{Bond-Lamberty}
\Author[4]{Susan}{Cook-Patton}
\Author[5]{Sandro}{Federici}
\Author[6]{Helene}{Muller-Landau}
\Author[7]{Camille}{Piponiot}
\Author[1]{Teagan}{Rogers}
\Author[5]{Valentyna}{Slivinska}
\Author[1,6 *]{Kristina J.}{Anderson-Teixeira}

\affil[1]{Center for Conservatiton Ecology, Smithsonian Conservation
Biology Institute, Front Royal, VA, United States}
\affil[2]{School of Geography, University of Leeds, Leeds, UK}
\affil[3]{Joint Global Change Research Institute, Pacific Northwest
National Laboratory, College Park, MD, United States}
\affil[4]{The Nature Conservancy; Arlington VA 22203, USA}
\affil[5]{IPCC Task Force on National Greenhouse Gas Inventories
Technical Support Unit, Institute for Global Environmental Strategies,
Hayama, Japan}
\affil[6]{Forest Global Earth Observatory, Smithsonian Tropical Research
Institute, Panama, Republic of Panama}
\affil[7]{CIRAD, Montpellier, France}

@@

\textbf{As we seek to mitigate climate change
\citep{unfccc_adoption_2015}, forests are critical.} \emph{(See intros
to other papers for relevant refs)} Under Paris Agreement,
\textbf{\#\#}\% of Nationally Determined Contributions relate to forests
\citep{grassi_key_2017}.

\textbf{The International Panel on Climate Change (IPCC) provides
guidance for national greenhouse gas inventories \citep[(REFS for older
guidelines),][]{ipcc_2019_2019}.} These include guidance for accounting
for forest land \citep{ipcc_agriculture_2006, ipcc_2019_2019a}. This
guidance has improved as more of the relevant underlying data has become
available \citep{requenasuarez_estimating_2019}, but there remains room
for continuous improvement as the science advances. For example,
\citet{cook-patton_mapping_2020} \emph{(summarize IPCC comparison
results)}. \emph{Moreover, it is useful for those compiling national
greenhouse gas inventories to have access to locally-specific
information, when available.} To improve the data accessible for C
accounting, the IPCC created the Emission Factor Database (EFDB;
https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a
recognized library of emission factors and other parameters that can be
used for estimating greenhouse gas emissions and removals.

\textbf{The Global Forest Carbon Database, ForC, is the largest
collection of published estimates of forest carbon stocks, increments,
and annual fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}.}
\emph{(add stats/ details)} As such, ForC is positioned to improve
forest C accounting through the transfer of data to EFDB. The purpose of
this publication is to document that process and provide recommendations
for future improvements.

\textbf{Here, we} (1) clarify definitions of relevant carbon stocks and
increments (2) describe mapping of ForC to IPCC's EFDB, (3) describe
updates to ForC (ForC v4.0), (4) summarize the data in ForC that's
relevant to EFDB, identifying gaps, and (5) provide recommendations for
improving data collection, analysis, database, and accounting.
